---
title: "AAC Import/clean"
output:
  html_document:
    df_print: paged
---

## Goals

- Are animal intakes increasing or decreasing each year since 2016? Plot a column chart of intakes by year. Don’t include 2021 since it is not a full year.
- Are there seasonal monthly trends in overall animal intakes? Plot intakes by year/month from 2016 to current, including 2021. (One long line or column chart)
- Plot that same monthly trend, except with month as the x axis, with a new line for each year.
- Do certain types of animals drive seasonal trends? Use five full years of data (not 2021) to summarize intakes by animal type and month. Plot with month on the x axis with a line for each animal type.

## Setup

```{r setup}
library(tidyverse)
library(janitor)
library(lubridate)
```

## Import the data

```{r import}
raw_data <- read_csv("data-raw/Austin_Animal_Center_Intakes.csv") %>% clean_names()

raw_data %>% head()
```

Glimpse the data

```{r glimpse}
raw_data %>% glimpse()
```

## Fix the dates

```{r date-fix}
date_fix <- raw_data %>% 
  mutate(
    timestamp = mdy_hms(date_time),
    intake_date = str_sub(date_time, 1, 10),
    intake_date = mdy(intake_date)
  )

date_fix %>% glimpse()
```

## Parse the date into helpful variables

Making year, month, and year-month columns to help plot later

```{r parse}
date_parts <- date_fix %>% 
  mutate(
    yr = year(intake_date), # creates yr and fills it with YYYY
    mo = month(intake_date, label = TRUE), # used label to write out month
    yr_mo = format(intake_date, "%Y-%m") # new rule to make YYYY-MM
  ) %>% 
  select(intake_date, yr, mo, yr_mo, everything(), -date_time, -month_year)

# peek
date_parts %>% glimpse()
```

## Filter dates

Filtering only 2016 to present, excluding partial month of October. Use through Sept. 2021

```{r filter-dates}
filtered_dates <- date_parts %>% 
  filter(intake_date >= "2016-01-01",
         intake_date <= "2021-09-30")

filtered_dates %>% summary()
```

## Export the data

```{r export}
filtered_dates %>% write_rds("data-processed/01-intakes.rds")
```








