---
title: "AAC Analysis"
output:
  html_document:
    df_print: paged
---

## Setup

```{r setup}
library(tidyverse)
library(lubridate)
library(scales)
library(plotly)
library(clipr)
```

## Import the data

Loading data from previous notebook

```{r import}
intakes <- read_rds("data-processed/01-intakes.rds")

intakes %>% glimpse()
```

## Question 1: Intakes by year

Finding whether animal intakes are increasing or decreasing each year since 2016. Plotting column chart of intakes by year, excluding 2021.

```{r q1-data}
intakes_yr_data <- intakes %>% 
  count(yr, name = "count_animals") %>% 
  filter(yr < "2021")

intakes_yr_data
```

Plot the data as a column chart

```{r q1-plot}
ggplot(intakes_yr_data, aes(x = yr, y = count_animals)) +
  geom_col() +
  geom_text(aes(label = comma(count_animals)), vjust = 2, color = "white") +
  scale_y_continuous(label = comma) +
  labs(
    title = "Austin Animal Center Intakes Since 2016",
    subtitle = "Data from Austin Animal Center, excluding 2021 data",
    caption = "By Cecilia Garzella",
    x = "Year",
    y = "Number of Animal Intakes"
  )
```

## Question 2: Animal intakes by month

```{r yrm-data}
intakes_yrmo <- intakes %>% 
  count(yr_mo, name = "count_animals")

# peek
intakes_yrmo
```

Plot the data as a line

```{r yrm-plot, fig.width = 7}
intakes_yrmo %>% 
  ggplot(aes(x = yr_mo, y = count_animals)) +
  geom_line(group = 1) +
  theme(axis.text.x = element_text(angle = 90)) # flips the x-axis name
```

## Question 3: Plot by month, split by year

```{r mo-yr-data}
mo_yr_data <- intakes %>% 
  count(mo, yr, name = "count_animals")

# peek
mo_yr_data
```

Plot by month, split by year

```{r mo-yr-plot}
mo_yr_data %>% 
  ggplot(aes(x = mo, y = count_animals)) +
  geom_line(aes(group = yr, color = as.character(yr))) +
  labs(
    title = "Austin Animal Center Intakes by Month, Split by Year",
    subtitle = "Data from Austin Animal Center, 2016-2021",
    caption = "By Cecilia Garzella",
    x = "Month",
    y = "Number of Intakes",
    color = "Year"
  )
```

## Question 4: Animal types by month

### Recoding the animal type

```{r types-data}
intake_types <- intakes %>% 
  filter(yr < 2021) %>% # removes 2021 because partial year
  mutate(
    animal_type = recode(animal_type, # first argument is column to look at
          "Bird" = "Other", # old name = new name
          "Livestock" = "Other"
    )
  )

intake_types %>% 
  count(animal_type)
```

### Prep the data for plotting

Creating animal type summary by month

```{r prep}
mo_types_data <- intake_types %>% 
  count(mo, animal_type, name = "count_type")

# peek
mo_types_data
```

### Plotting animal types by month

Splitting by animal type

```{r types-plot}
mo_types_data %>% 
  ggplot(aes(x = mo, y = count_type)) +
  geom_line(aes(group = animal_type, color = animal_type)) +
  labs(
    title = "Animal Types Driving Seasonal Trends at Austin Animal Center",
    subtitle = "Monthly intake counts contain data collected over 5-year period (2016-2020)",
    caption = "Data from Austin Animal Center",
    x = "Month",
    y = "Number of Intakes",
    color = "Animal Type"
  )
```

I was surprised to see that intakes of dogs and other animals were pretty consistent year-round, but cats were driving the spring/summer surge. I did a quick Google search and read that cats generally reproduce around this time, peaking in April. This could be one possible explanation, but to find out more information, I would do some reporting and talk to the folks at AAC.



